Out-of-Sample Comparisons of Overfit Models

### 2014-03-25 paper.tex

This section presents an analysis of equity premium predictability
similar to \citet{GoW:08}.%
\footnote{Calculations in this section are done in R \citep{Rde:10}
  using the lmtest \citep{ZeH:02} and sandwich \citep{Zei:04}
  packages. Tables and grapsh were prepared with the Hmsic
  \citep{Har:10} and tikzDevice \citep{ShB:11:0.6.1} R packages as
  well as the TikZ and PGF LaTeX packages \citep{Tan:10}} %
We estimate the expected forecasting
performance of the largest model they consider, a model with 13
regressors, using 81 observations (annual data from 1928 to 2009).  We
construct confidence intervals for the difference in performance
between this model and a prevailing mean benchmark, and do so for
different values of $R$ and $P$ to examine the effect of the training
and test sample choice on this OOS statistic.  We find that the
estimate of the larger model's Mean Squared Error (MSE) decreases relative to the
benchmark as $R$ increases, but it is never smaller than the
benchmark's.  This result, combined with the larger model's in-sample
significance, suggests that there may be a real relationship between
the equity premium and some of these predictors, but that the
relationship can not be estimated accurately enough to be useful for
forecasting and supports \citepos{GoW:08} conclusions.  These results
hold even when imposing \citepos{CaT:08} restriction that the equity
premium forecast be non-negative.

We'll start with a very brief review.  \citet{GoW:08}
study the OOS forecasting performance of different variables thought
to predict the equity premium (calculated as the difference between
the return on the S\&P 500 index and the T-bill rate).  Some of these
variables are listed in
Table~\ref{tab:equity}.%
\footnote{Table~\ref{tab:equity} only lists the
  variables used in \citepos{GoW:08} ``kitchen sink''
  model.  Some of the variables that they use in bivariate models are
  excluded from this model either because the series are too short or
  because the variables are linear combinations of other variables.} %
\citet{GoW:08} estimate the OOS forecasting
performance of bivariate models of the form
\[
r_{t+1} = \beta_0 + \beta_1 x_t + \e_{t+1}
\]
using OLS with a recursive window, where $r_{t+1}$ is the equity
premium and $x_t$ is a generic predictor, as well as some larger
models.  They find that these models forecast no better than the
prevailing mean of the equity premium.
\citet{CaT:08} find that nonlinear extensions of
Goyal and Welch's models are more accurate, either imposing sign
restrictions on the estimated coefficients or imposing that the
forecast be non-negative, but do not test for significance.

Obviously, \citepos{GoW:08} bivariate models do not
match the asymptotic theory of this paper, so we restrict our
attention here to their ``kitchen sink'' model, which includes all of
the variables, except for a few dropped because of data availability or
(perfect) multicollinearity.  Formally, the benchmark and alternative models are
\begin{eqnarray}
  \label{eq:3}
  r_{t+1} &=& \mu + \e_{1,t+1} \\
  \label{eq:2}
r_{t+1} &=& \beta_0 + \sum_{i=1}^K \beta_{i} x_{it} + \e_{2,t+1}
\end{eqnarray}
respectively and are estimated by OLS using the fixed-window scheme.
The predictors are listed in Table~\ref{tab:equity}; for a detailed
description of each predictor, please see Goyal and Welch's original
paper.  In this model, there are 13 regressors (including the constant
term) and we estimate the coefficients using annual data from 1928 to
2009.  This makes the ratio $K/T$ equal to about 0.16.  We also
present results for one of \citepos{CaT:08}
restricted forecasting models.  For this model, we estimate
\eqref{eq:3} and \eqref{eq:2}, but impose that $\hat r_{t+1}$ be
non-negative for each forecast.  We then calculate the OOS test just
as for the unconstrained model.

Table~\ref{tab:gwinsample} gives the results of the full sample
regression for Equation~\eqref{eq:2}, using Newey-West standard errors
with two lags \citep{NeW:87}.%
\footnote{We compare the test statistic
  to critical values from the $F$-distribution.} %
The $p$-value is
very small (less than 0.01), indicating that the coefficients are
nonzero in population and at least one of these predictors is
correlated with the equity premium.  As we argue earlier, this result
does not imply that the model will forecast well, which we look at
next.

To study the effect of the training sample size on the DMW statistic, we
calculate the one-sided confidence interval for $\E_T \oosB$ given
by \eqref{interval-greater} corresponding to the null and alternative
hypotheses
\[ H_0: \quad E_T \oosB \leq 0 \qquad
H_A: \quad E_T \oosB > 0
\]
using the fixed-window scheme for each value of $R$ between 20 and
$T-10$.  The standard deviation is estimated using a Newey-West
estimator with $\lfloor P^{1/4}\rfloor$ lags.  For small values of
$R$, the OOS average is expected to underestimate the performance of
the larger model relative to the smaller, but this may not hold in
this particular dataset.

Figures~\ref{fig:empirics1} plots the OLS results and
Figure~\ref{fig:empirics2} imposes \citepos{CaT:08} restriction.  The
solid line in each figure shows the OOS average, $\oosA$, and the
shaded region indicates the 95\% one-sided confidence interval implied
by the DMW test.  Negative numbers indicate that the kitchen sink
model has higher out-of-sample loss.  We can see that the same
patterns hold for both models: the performance difference decreases as
$R$ grows, but the kitchen sink model is never more accurate.  We also
see that the performance difference decreases suddenly over the period
$R=29$ to $R=34$ (corresponding to the years 1956--1961).
Figure~\ref{fig:empirics3} plots the accuracy of the individual
forecasts (only for the linear models) and shows that this change is
the result of a sudden improvement in the kitchen sink model.  This
change may indicate instability in the underlying relationship, as
\citet{GoW:08} propose.

In summary, we fail to reject the null that the benchmark prevailing
mean model is more accurate than \citepos{GoW:08} kitchen sink.  This
result is consistent with Goyal and Welch's original analysis.  Unlike
Goyal and Welch, we attribute this result, at least in part, to
parameter uncertainty---the full sample results indicate that the
larger model could predict better than the benchmark with enough
data.%
\footnote{\citet{BWB:10} make a similar point about exchange rate
models, but see also \citet{Chi:10} and \citet{Gia:10}.} %
Obviously, this is an illustrative exercise only and is not meant to
be comprehensive.  In this dataset, there is likely parameter
instability that we have not addressed, and we have not dealt with
variable persistence at all.  However, as we mention earlier, our
requirement that $P$ be small should let our theoretical results
extend to moderate instability, and the fact that we do not rely on
the asymptotic distribution of the coefficient estimates should
mitigate the effect of persistence as well.

@@

\begin{table}[tb]
  \begin{tabularx}{\linewidth}{XX}
    \toprule
    Category & Variable \\
    \midrule
    Stock market variables
    & Dividend Price ratio (log) \\
    & Earnings Price ratio (log) \\
    & Stock Market Variance \\
    & Book to Market Ratio \\
    & Net Equity Expansion \\
    & Percent Equity Issuing \\\\
    Interest rate variables
    & Treasury Bill rate (3 month) \\
    & Long Term Yield \\
    & Long Term Rate \\
    & Default Return Spread \\
    & Default Yield Spread \\
    & Inflation \\
    \bottomrule
  \end{tabularx}
  \caption{Variables used to predict the equity premium.
    Please see Goyal and Welch's original paper \citep{GoW:08}
    for a detailed description of each variable.}
  \label{tab:equity}
\end{table}

\begin{table}[b]
 \begin{center}
 \begin{tabular}{lrrrr}\hline\hline
   \multicolumn{1}{l}{waldtest}&\multicolumn{1}{c}{Res.Df}&\multicolumn{1}{c}{Df}&\multicolumn{1}{c}{F}&\multicolumn{1}{c}{Pr(\textgreater
     F)}\tabularnewline
   \hline
   1&69&&&\\
   2&81&$-$12&3.4&0.00066 \\
   \hline
\end{tabular}
\caption{Wald test for the null hypothesis that all of the
  coefficients in Goyal and Welch's (2008) ``kitchen sink'' model are
  zero, except for the intercept, using the Newey-West
  variance-covariance matrix.  The p-value is calculated from the $F$
  distribution.\label{tab:gwinsample}}
\end{center}
\end{table}

>>>>

This section presents a study of equity premium predictability based
on \citet{GoW:08}. Goyal and Welch conduct an out of sample analysis
of 18 different variables thought, based on previous research, to
predict the equity premium (calculated as the difference between the
return on the S\&P 500 index and the T-bill rate).%
\footnote{\citet{GoW:08} builds on previous research by \citet{BoH:99}
  and \cite{GoW:03}.} %
Their analysis has two notable features: they compare different
predictors across the same time periods and frequency as much as
possible, making the accuracy of these predictors directly comparable;
many of the papers that originally proposed these predictors used
different time periods, methods, or observational frequencies so their
results were not directly comparable. And Goyal and Welch compare
these predictors out-of-sample; many of the original studies found
in-sample evidence of equity premium predictability but did not look
at out-of-sample evidence.

Goyal and Welch consider many different models; most of them are very
simple---regression onto a constant and a single stochastic
predictor---but some are more complicated. They find that essentially
none of the models outperform a simple benchmark, the prevailing mean
of the equity premium, and conclude that these variables do not
predict the equity premium. A large literature has subsequently sought
to explain or rebut these results, including several responses to
Goyal and Welch's original working paper published in the same special
issue of \textit{The Review of Financial Studies} as \citet{GoW:08}.%
\footnote{Those papers are \citet{CaT:08}, \citet{Coc:08},
  \citet{BRW:08}, and \citet{LeN:08}.} %

\citet{GoW:08}, as well as \citet{BoH:99}, \citet{LeN:08}, and many
other authors, propose that instability could explain the OOS failure
of these models. However, we have shown in this paper that overfit can
also explain this pattern, significant in-sample results that do not
hold up out of sample. In this section, we explore the extent to which
overfit is a potential concern in this data set and estimate the
expected forecasting performance of the largest model they consider, a
model with 13 regressors, using 81 observations (annual data from 1928
to 2009). The predictors are listed in Table~\ref{tab:equity}; please
see Goyal and Welch's original paper for detailed information about
these variables.%
\footnote{Table~\ref{tab:equity} only lists the variables used in
  \citepos{GoW:08} ``kitchen sink'' model.  Some of the variables that
  they use are excluded from this model either
  because the series are too short or because the variables are linear
  combinations of other variables.} %

\citet{GoW:08} focus primarily on univariate regression models of the
form
\[
r_{t+1} = \beta_0 + \beta_1 x_{it} + \e_{t+1}
\]
using OLS with a recursive window, where $r_{t+1}$ is the equity
premium and $x_{it}$ is one of the predictors listed in
Table~\ref{tab:equity}.%
\footnote{To make it easier to compare coefficients across variables
  and models, all of the regressors are studentized before our
  analysis and the equity premium is reported in basis points.} %
This focus is consistent with much of the rest of the literature. But
it is not clear that this approach---using a large number of
restricted models---is any more reliable than using a single, large
model. Very few papers in the equity-premium predictability literature
explicitly account for the multiplicity of model comparisons%
\footnote{\citet{RaW:06}, \citet{RaZ:12}, and \citet{Cal:13b} are
  exceptions.} %
and most widely-used statistics for multiple comparisons are
derived under the assumption that there is a finite number of
hypotheses or models (as is the case in \citealp{STW:99c},
\citealp{Whi:00}, \citealp{Han:05}, and \citealp{LeR:05}), so there is
little theoretical evidence that they are any more reliable than a
regression model that encompasses all of the smaller models. Moreover,
since \citet{GoW:08} (along with most of the rest of the literature)
want to interpret these univariate regressions as meaningful
statements about the true relationships between the equity premium and
the regressors, omitted variable bias is a serious potential issue.

For example, Table~\ref{tab:gwinsample} presents two full-sample
coefficient estimates for each of the 12 predictors we consider.%
\footnote{For all of our full sample results, we studentize the
  regressors to make it easier to compare coefficient estimates and we
  express the equity premium in basis points.} %
The first column is the estimator of the coefficient on the listed
variable in the full regression,
\begin{equation}\label{eq:30}
  r_{t+1} = \beta_0 + \sum_{i=1}^{12} \beta_{i} x_{it} + \e_{t+1},
\end{equation}
and the third column is the $p$-value associated with a two-sided
$t$-test that the coefficient is zero in population. The second column
is the estimator of the coefficient from the univariate regression
\begin{equation}
  r_{t+1} = \alpha_i + \gamma_{i} x_{it} + \e_{t+1},
\end{equation}
and the fourth column is the $p$-value associated with its $t$-test.%
\footnote{All of the standard errors were calculated using a
  Newey-West kernel with two lags.} %
Some of the coefficient estimates agree, but some do not. The
coefficient estimate for the \emph{default yield spread}, for example,
is substantially and significantly negative in the full model ($-7.79$
with $p$-value $0.042$), but is near zero in the univariate regression
($0.79$ with $p$-value $0.720$). This pair of results implies that the
the default yield spread contains information about the equity premium
but is also correlated with other poor predictors, and that this
additional correlation adds noise to the univariate regression
and masks the true relationship. Similarly, \emph{net equity expansion}
is significant at 10\% in the univariate regression but not in the
full model ($p$-values of $0.060$ and $0.626$ respectively), which is
likely attributable to omitted variable bias. So there is merit to
studying a large model that includes all of the variables.

Results for the full sample estimates of Equation~\eqref{eq:2} are in
Table~\ref{tab:gwinsample}.%
\footnote{Calculations in this section are done in R \citep{Rde:10}
  using the \emph{lmtest} \citep{ZeH:02} and \emph{sandwich}
  \citep{Zei:04} packages. Tables and graphs were prepared with the
  \emph{dbframe} \citep{Cal:10b}, \emph{Hmsic} \citep{Har:10},
  \emph{tikzDevice} \citep{ShB:11:0.6.1}, and \emph{xtable}
  \citep{Dah:09} R packages as well as the \emph{booktabs}
  \citep{FeE:05} and \emph{TikZ and PGF} LaTeX packages
  \citep{Tan:10}}%
\footnote{We compare the test statistic to critical values from the
  $F$-distribution, which has been shown to be more reliable when
  there are many regressors \citep{Ana:12,Cal:11c}.} %
The last rows of the table list measures of the full-model fit.  The
$p$-value for the test of full model fit is very small (less than
0.01), indicating that the coefficients are nonzero in population and
at least one of these predictors is correlated with the equity
premium.%
\footnote{We have done additional analysis to try to identify which
  predictors were correlated with the equity premium, but none of the
  individual regressors were significant after correcting for
  multiplicity. The Bonferroni correction, for example, suggests that
  an individual $p$-value would need to be less than $0.10/12 \approx
  0.0083$ for its corresponding coefficint to be significant at the
  10\% level, but the smallest $p$-value is $0.035$. One can improve
  on the Bonferroni correction, but a comprehensive analysis is beyond
  the scope of this paper.} %
As we argue throughout the paper, this result does not imply that the
model will forecast well.

To determine whether the full model can forecast well, we use the DMW
\oost\ test to compare it to a sample mean benchmark,
\begin{equation}\label{eq:2}
r_{t+1} = \mu + \e_{1,t+1}.
\end{equation}
Both models are estimated by OLS using the fixed-window scheme. We
also present results for a restricted model proposed by
\citet{CaT:08} that imposes that $\hat r_{t+1}$ be non-negative for
each forecast.
To study the effect of the training sample size on the DMW statistic, we
calculate the one-sided confidence interval for $\E_T \oosB$ given
by \eqref{interval-greater} corresponding to the null and alternative
hypotheses
\[ H_0: \quad E_T \oosB \leq 0 \qquad
H_A: \quad E_T \oosB > 0
\]
using the fixed-window scheme for each value of $R$ between 20 and
$T-10$. The standard deviation is estimated using a Newey-West
estimator with $\lfloor P^{1/4}\rfloor$ lags.  For small values of
$R$, the OOS average is expected to underestimate the performance of
the larger model relative to the smaller, but this may not hold in
this particular dataset. For the OOS results, we express the equity
premium in perecentage points.

Figure~\ref{fig:empirics1} plots the OLS results and
Figure~\ref{fig:empirics2} imposes \citepos{CaT:08} restriction. The
solid line in each figure shows the OOS average, $\oosA$, and the
shaded region indicates the 95\% one-sided confidence interval implied
by the DMW test.  Negative numbers indicate that the full model has
higher out-of-sample loss.  We can see that the same patterns hold for
both models: the performance difference decreases as $R$ grows, but
the full model is never more accurate.  We also see that the
performance difference decreases suddenly over the period $R=29$ to
$R=34$ (corresponding to the years 1956--1961).
Figure~\ref{fig:empirics3} plots the accuracy of the individual
forecasts (only for the linear models) and shows that this change is
the result of a sudden improvement in the full model.  This change may
indicate instability in the underlying relationship, as proposed by
\citet{GoW:08}.

In summary, we fail to reject the null that the benchmark prevailing
mean model is more accurate than the full model including all of
\citepos{GoW:08} predictors.  This result is consistent with Goyal and
Welch's original analysis.  Unlike Goyal and Welch, we attribute this
result, at least in part, to parameter uncertainty---the full sample
results indicate that there is a true predictive relationship between
some of these variables and the equity premium and the larger model
could predict better than the benchmark with enough data.%
\footnote{\citet{BWB:10} make a similar point about exchange rate
  models, but see also \citet{Chi:10} and \citet{Gia:10}.} %
These also indicate that combination or shrinkage estimators of the
full model could significantly improve on the benchmark.%
\footnote{See \citet{RaZ:12} for a recent review of this
  literature.} %

@@

\begin{table}[tb]
  \begin{tabularx}{\linewidth}{XX}
    \toprule
    Category & Variable \\
    \midrule
    Stock market variables
    & Dividend to price ratio (log) \\
    & Earnings to price ratio (log) \\
    & Stock market variance \\
    & Book to market ratio \\
    & Net equity expansion \\
    & Percent equity issuing \\\\
    Interest rate variables
    & Treasury Bill rate (3 month) \\
    & Long term yield \\
    & Long term rate \\
    & Default return spread \\
    & Default yield spread \\
    & Inflation rate \\
    \bottomrule
  \end{tabularx}
  \caption{Variables used to predict the equity premium
    (Section~\ref{sec:empirics}).
    Please see Goyal and Welch's original paper \citep{GoW:08}
    for a detailed description of each variable.}
  \label{tab:equity}
\end{table}
\clearpage

\tryinput{floats/empirics-insample.tex}
\tryinput{floats/empirics-insample-tuned.tex}
% These macros create tables. They are defined in
% floats/empirics-insample.tex and possibly overwritten in the next
% file.
\insamplecoefs
\insampletests
